Supplement: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
fixed weird citation bug

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.

>>>>

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG),
for which we have effective approaches like the Inside/Outside algorithm~(Lafferty,~2000), %\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.

### 2018-02-09 supplement.tex
fix latex errors
change listings style for consistency
fixed weird citation bug

\usepackage{amssymb}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{tabularx}

@@

Differentiating with respect to $\theta_e$ and setting to zero:
%% \begin{align}
%%   & \frac{\alpha + \sum_x\expect_{Q_x}\left[ c(e,p)\right]}{\theta_e} -
%% \theta_e\sum_x  \expect_{Q_x}\left[c(p)\frac{r(p,e)}{\sum_{e'\in \mathcal{D}} r(e',p)\theta_{e'}} \right]  
%%   \end{align}
%%
\begin{align*}
  \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}} \\
  \frac{c(x,p)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
\end{align}
where $N=c(p)$ and $a(x) = r(x, p)$. From here forward, we similarly let
$c(x)=c(x,p)$ for notational convenience.

@@

\section{Example Program Induction Tasks}

\subsection{String Editing}

@@

Here we present representative DSLs learned by
our model. DSL primitives discovered by the algorithm are prefixed with \code{#}.
Variables are prefixed with \$, and we use De Bruijn indices to model variables~\cite{pierce}.
We abbreviate lambda as \textbackslash.

@@

\begin{verbatim}
#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2))))))
#(\ (\ (nand (nand $0 $1) (nand $0 $1))))
#(\ (\ (nand (nand $0 $0) (nand $1 $1))))
#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 $3))))))
#(\ (\ (\ (nand (nand $0 $0) (nand $1 (nand $2 $1))))))
#(\ (\ (#(\ (\ (nand (nand $0 $1) (nand $0 $1)))) $0 (nand $1 $0))))
#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0))))))
#(\ (\ (#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0)))))) $0 $1 (nand $1 $1))))
#(\ (\ (nand $0 (nand $1 $0))))
#(\ (#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2)))))) (nand $0 $0)))
#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $3 $2)))))))
\end{verbatim}

\subsection{Symbolic regression}
\begin{verbatim}
#(\ (* (+ REAL $0) REAL))
#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0)))
#(\ (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))
#(\ (#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0))) (+ $0 $0)))
#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0)))
#(\ (* (#(\ (+ $0 REAL)) $0) (#(\ (+ (* $0 (* $0 REAL)) REAL)) $0)))
#(\ (+ (* $0 (* $0 REAL)) REAL))
#(* REAL)
#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0))
#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0))))
#(\ (+ $0 REAL))
\end{verbatim}

\subsection{String editing}
\begin{verbatim}
#(\ (\ (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0)))))
#(\ (\ (join (chr->str $0) (map (\ (strip $0)) $1))))
#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0))))
#(\ (#(join empty) (split '.' $0)))
#(#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2))))) (decr 0))
#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (slice $2 (incr $2) $0)))))
#(#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (incr 0))
#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2)))))
#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1))))))
#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr $0)))
#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (slice $0 (incr (incr $0)))) $2 $0)))))
#(\ (\ (slice $0 (incr $0) (nth (incr $0) $1))))
#(\ (\ (#(\ (\ (slice $0 (len $1) $1))) (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1) (incr 0))))
#(\ (\ (caseUpper (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0))))))
#(\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (nth 0 $1)) $0)))
#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr (decr $0))))
#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1))))
#(\ (\ (#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0)))) (split '.' (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1)) '.')))
#(\ (\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1))))) $1 $2 0)) $0))))
#(\ (\ (slice $0 (incr (incr $0)) (#(nth (incr 0)) $1))))
#(join empty)
#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1)))))
#(\ (\ (nth 0 (split $0 (#(nth (incr 0)) $1)))))
#(\ (slice $0 (incr (incr $0))))
#(nth (incr 0))
#(\ (join empty (map (\ (caseCapitalize $0)) $0)))
#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2)))))
#(\ (\ (slice $0 (len $1) $1)))
#(\ (\ (\ (join (chr->str $0) (map $1 $2)))))
#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1))))
\end{verbatim}

\subsection{List functions}
\begin{verbatim}
#(\ (\ (index $0 (sort $1))))
#(\ (\ (slice $0 (+ $1 4))))
#(mapi (\ (\ (eq? $0 $2))))
#(\ (mapi (\ (\ (gt? $0 $2)))))
#(\ (any (\ (eq? $0 $1))))
#(\ (mapi (\ (\ (+ $0 $2)))))
#(\ (\ (mapi (\ (\ (index $2 $3))) $1)))
#(\ (mapi (\ (\ (mod $0 $2)))))
#(\ (mapi (\ (\ (* $0 $2)))))
#(\ (\ (singleton (index $0 $1))))
#(\ (\ (slice $0 (sum $1) $1)))
#(slice 0)
#(\ (\ (index (negate $0) (sort $1))))
#(\ (++ (singleton $0)))
#(\ (\ (++ $0 (singleton $1))))
#(\ (\ (++ $0 (singleton (index $1 $0)))))
#(\ (\ (++ (singleton (index $0 $1)) $1)))
#(\ (\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) $1 $2)))))
#(\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))
#(#(\ (mapi (\ (\ (gt? $0 $2))))) 3 empty)
#(all (\ (is-square $0)))
#(\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) 1 $1))))
#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1))))
#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2)))
#(filter (\ (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0))))
#(\ (\ (\ (++ (#(slice 0) $0 $1) (#(\ (\ (slice $0 (sum $1) $1))) $1 $2)))))
#(\ (filter (\ (#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1)))) $1 (singleton $0)))))
#(filter (\ (not (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0)))))
#(\ (++ $0 (++ $0 $0)))
#(\ (++ $0 (#(\ (++ $0 (++ $0 $0))) $0)))
#(index 0)

\end{verbatim}

>>>>

\usepackage{amssymb}
\usepackage{listings}
\lstset{%
  basicstyle=\footnotesize\ttfamily,
  breaklines=true}
\usepackage{wrapfig}
\usepackage{tabularx}

@@

Differentiating with respect to $\theta_e$ and setting to zero:
%% \begin{align}
%%   & \frac{\alpha + \sum_x\expect_{Q_x}\left[ c(e,p)\right]}{\theta_e} -
%% \theta_e\sum_x  \expect_{Q_x}\left[c(p)\frac{r(p,e)}{\sum_{e'\in \mathcal{D}} r(e',p)\theta_{e'}} \right]  
%%   \end{align}
%%
\begin{align*}
  \frac{c(e, p)}{\theta_e} &= c(p)\frac{r(e, p)}{\sum_{e'\in\mathcal{D}} r(e', p)\theta_{e'}} \\
  \frac{c(x,p)}{\theta_x} &= N\frac{a(x)}{\sum_y a(y)\theta_y}
\end{align*}
where $N=c(p)$ and $a(x) = r(x, p)$. From here forward, we similarly let
$c(x)=c(x,p)$ for notational convenience.

@@

\section{Example String Editing Tasks}

@@

Here we present representative DSLs learned by
our model. DSL primitives discovered by the algorithm are prefixed with
\lstinline!#!.
Variables are prefixed with \lstinline!$!, and we use De Bruijn indices to model variables~\cite{pierce}.
We abbreviate lambda as \lstinline!\!.

@@

\begin{lstlisting}
#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2))))))
#(\ (\ (nand (nand $0 $1) (nand $0 $1))))
#(\ (\ (nand (nand $0 $0) (nand $1 $1))))
#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 $3))))))
#(\ (\ (\ (nand (nand $0 $0) (nand $1 (nand $2 $1))))))
#(\ (\ (#(\ (\ (nand (nand $0 $1) (nand $0 $1)))) $0 (nand $1 $0))))
#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0))))))
#(\ (\ (#(\ (\ (\ (nand (nand $0 $1) (nand $0 (nand $2 $0)))))) $0 $1 (nand $1 $1))))
#(\ (\ (nand $0 (nand $1 $0))))
#(\ (#(\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $0 $2)))))) (nand $0 $0)))
#(\ (\ (\ (\ (nand (nand $0 $1) (nand $2 (nand $3 $2)))))))
\end{lstlisting}

\subsection{Symbolic regression}
\begin{lstlisting}
#(\ (* (+ REAL $0) REAL))
#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0)))
#(\ (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))
#(\ (#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0))) (+ $0 $0)))
#(\ (#(\ (+ $0 REAL)) (#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0)) $0)))
#(\ (* (#(\ (+ $0 REAL)) $0) (#(\ (+ (* $0 (* $0 REAL)) REAL)) $0)))
#(\ (+ (* $0 (* $0 REAL)) REAL))
#(* REAL)
#(\ (* (#(\ (#(\ (+ $0 REAL)) (#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0)))) $0))) $0) $0))
#(\ (* $0 (#(\ (+ (* $0 (* $0 REAL)) REAL)) (#(\ (+ $0 REAL)) $0))))
#(\ (+ $0 REAL))
\end{lstlisting}

\subsection{String editing}
\begin{lstlisting}
#(\ (\ (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0)))))
#(\ (\ (join (chr->str $0) (map (\ (strip $0)) $1))))
#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0))))
#(\ (#(join empty) (split '.' $0)))
#(#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2))))) (decr 0))
#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (slice $2 (incr $2) $0)))))
#(#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (incr 0))
#(\ (\ (\ (#(\ (\ (slice $0 (len $1) $1))) (nth $0 $1) (decr $2)))))
#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1))))))
#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr $0)))
#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (slice $0 (incr (incr $0)))) $2 $0)))))
#(\ (\ (slice $0 (incr $0) (nth (incr $0) $1))))
#(\ (\ (#(\ (\ (slice $0 (len $1) $1))) (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1) (incr 0))))
#(\ (\ (caseUpper (#(nth (incr 0)) (split $0 (#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1)))) $1 $0))))))
#(\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (nth 0 $1)) $0)))
#(\ (#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2))))) (decr (decr $0))))
#(\ (\ (join (chr->str $0) (map (\ (caseCapitalize $0)) $1))))
#(\ (\ (#(\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (++ $0 $0)))) (split '.' (#(\ (\ (strip (nth 0 (split $0 (#(nth (incr 0)) $1)))))) $0 $1)) '.')))
#(\ (\ (\ (#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1)))) (split $0 (#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1))))) $1 $2 0)) $0))))
#(\ (\ (slice $0 (incr (incr $0)) (#(nth (incr 0)) $1))))
#(join empty)
#(\ (\ (\ (++ (nth $0 (split $1 $2)) (chr->str $1)))))
#(\ (\ (nth 0 (split $0 (#(nth (incr 0)) $1)))))
#(\ (slice $0 (incr (incr $0))))
#(nth (incr 0))
#(\ (join empty (map (\ (caseCapitalize $0)) $0)))
#(\ (\ (#(\ (\ (\ (join (chr->str $0) (map $1 $2))))) $0 (\ (#(\ (\ (slice $0 (len $1) $1))) $0 $2)))))
#(\ (\ (slice $0 (len $1) $1)))
#(\ (\ (\ (join (chr->str $0) (map $1 $2)))))
#(\ (\ (++ (chr->str $0) (#(nth (incr 0)) $1))))
\end{lstlisting}

\subsection{List functions}
\begin{lstlisting}
#(\ (\ (index $0 (sort $1))))
#(\ (\ (slice $0 (+ $1 4))))
#(mapi (\ (\ (eq? $0 $2))))
#(\ (mapi (\ (\ (gt? $0 $2)))))
#(\ (any (\ (eq? $0 $1))))
#(\ (mapi (\ (\ (+ $0 $2)))))
#(\ (\ (mapi (\ (\ (index $2 $3))) $1)))
#(\ (mapi (\ (\ (mod $0 $2)))))
#(\ (mapi (\ (\ (* $0 $2)))))
#(\ (\ (singleton (index $0 $1))))
#(\ (\ (slice $0 (sum $1) $1)))
#(slice 0)
#(\ (\ (index (negate $0) (sort $1))))
#(\ (++ (singleton $0)))
#(\ (\ (++ $0 (singleton $1))))
#(\ (\ (++ $0 (singleton (index $1 $0)))))
#(\ (\ (++ (singleton (index $0 $1)) $1)))
#(\ (\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) $1 $2)))))
#(\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))
#(#(\ (mapi (\ (\ (gt? $0 $2))))) 3 empty)
#(all (\ (is-square $0)))
#(\ (\ (#(\ (mapi (\ (\ (mod $0 $2))))) $0 (#(\ (mapi (\ (\ (+ $0 $2))))) 1 $1))))
#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1))))
#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2)))
#(filter (\ (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0))))
#(\ (\ (\ (++ (#(slice 0) $0 $1) (#(\ (\ (slice $0 (sum $1) $1))) $1 $2)))))
#(\ (filter (\ (#(\ (\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 $1)))) $1 (singleton $0)))))
#(filter (\ (not (#(\ (all (\ $0) (#(\ (\ (#(\ (mapi (\ (\ (eq? $0 $2))))) 0 (#(\ (mapi (\ (\ (mod $0 $2))))) $0 $1)))) $0 2))) (singleton $0)))))
#(\ (++ $0 (++ $0 $0)))
#(\ (++ $0 (#(\ (++ $0 (++ $0 $0))) $0)))
#(index 0)
\end{lstlisting}
